The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Work on section 4.2
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1807 4525493e-7705-40b1-a816-d608a930855b

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we are designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(Unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS core}
The DAS core communicated with different data-services and retrieved
data on demand. This was done by invoking appropriate data-service plug-in(s)
upon the user request. Each plug-in consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service plug-in contains a map about data-service APIs,
input and output parameters and their mapping to DAS keys. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
This information was stored into DAS Mapping DB which we discuss next.

Although the majority of data-service plug-ins looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required, e.g. access to data-services 
accessible via private network. Therefore, a way to propagate 
user credentials was required, leading to specific plug-in implementation.

%All user queries were written to Analytics DB for further analysis, while
%mapping between data services were kept in Mapping DB. The data-service 
%output were re-mapped into DAS notations and, if necessary, aggregated 
%inside of the DAS cache.

@@

%\bibitem{REST}
%Fielding, Roy Thomas ``Architectural Styles and the Design of 
%Network-based Software Architectures'', Doctoral dissertation, 2000,
%University of California, Irvine

>>>>

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we are designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. It conists of cacher server,
analytics and mapping databases. The data in a cache can be shared
across multiple queries. The information retrieval is triggered
on demand basis by invoking appropriate data-service APIs. The
optional DAS robots (Unix daemons) can be used to pre-fetch
the most populary query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS core}
The DAS core library was responsible for data retrieval on demand. 
It had the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
%For each GET request we performed the following action.
Each input query was recorded into Analytics DB for further 
analysis. It was decomposed into the set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requestor
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and store results into the cache. Finally, the results were
retrieved from the cache and delivered to the user.

The mapping between user queries, the DAS keys used in a query, 
and API names/parameters was stored in a Mapping DB. Such mapping was 
used by data-service plug-ins. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
represents the mapping between \verb+site+ DAS keyword and
\verb+CMSNametoSE+ API provided by SiteDB data-service (the data-service
URLs were kept in the configuration file).
The data-service plug-in implementation were similar and their logic was 
abstracted, while a few extentions were required, e.g. specific 
security modules for data-services accessible via private network.

@@

\bibitem{REST}
Fielding, Roy Thomas ``Architectural Styles and the Design of 
Network-based Software Architectures'', Doctoral dissertation, 2000,
University of California, Irvine
